---
title: "Reading Excel Workbooks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading Excel Workbooks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`fastexcel` reads Excel workbooks through a Rust backend and returns Arrow
objects by default. This keeps the read path efficient while still allowing you
to convert to common R data structures when needed.

```{r setup}
library(fastexcel)

file <- system.file("extdata/Pop_Density.xlsx", package = "fastexcel")
```

## Read a worksheet

Call `read_excel()` with a workbook path. The default return value is an
`arrow::Table`.

```{r read-default}
tbl <- read_excel(file)
tbl
```

You can select a sheet by 1-based index or by sheet name.

```{r read-sheet}
read_excel(file, sheet = 1, n_max = 2)
read_excel(file, sheet = "Sheet1", n_max = 2)
```

## Choose a return type

Use `as` when you want a specific R object type.

```{r return-types}
read_excel(file, as = "data.frame")
read_excel(file, as = "tibble")
read_excel(file, as = "arrow_record_batch")
```

For one-column reads, you can return either an Arrow array or a base R vector.

```{r one-column}
read_excel(file, range = "A:A", as = "arrow_array")
read_excel(file, range = "A:A", as = "vector")
```

## Select rows and columns

Use Excel-style column ranges for contiguous column selections.

```{r ranges}
read_excel(file, range = "A:C", as = "data.frame")
```

Use `columns` to select by column name or 1-based column position.

```{r columns}
read_excel(file, columns = c("city", "Year"), as = "data.frame")
read_excel(file, columns = c(1, 3), as = "data.frame")
```

Use `n_max` to limit the number of returned data rows.

```{r n-max}
read_excel(file, n_max = 3, as = "data.frame")
```

## Read workbook bytes

`file` may also be a raw vector containing workbook bytes. This is useful when a
cloud storage client returns the downloaded object directly in memory.

```{r raw-bytes}
bytes <- readBin(file, what = "raw", n = file.info(file)$size)
bytes[1:5]
read_excel(bytes, n_max = 2, as = "data.frame")
```

For example, `googleCloudStorageR::gcs_get_object()` can return object bytes as a
raw vector. Pass those bytes directly to `read_excel()` without writing a
temporary workbook to disk.

```{r gcs-bytes, eval = FALSE}
object <- "path/to/workbook.xlsx"

bytes <- googleCloudStorageR::gcs_get_object(object)
read_excel(bytes, as = "data.frame")
```

Keeping the workbook in memory avoids creating extra local file artifacts. That
is especially useful for sensitive data workloads where reducing disk writes can
lower the chance of leaving behind temporary files, cached copies, or cleanup
gaps outside the R process.
