---
title: "Workbook Metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workbook Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`fastexcel` includes helpers for inspecting a workbook before reading data. Use
these helpers to discover sheet names, dimensions, column metadata, tables, and
defined names.

```{r setup}
library(fastexcel)

file <- system.file("extdata/Pop_Density.xlsx", package = "fastexcel")
```

## Sheets

List available worksheets with `excel_sheets()`.

```{r sheets}
excel_sheets(file)
```

Use `excel_sheet_info()` to inspect worksheet dimensions and visibility.

```{r sheet-info}
excel_sheet_info(file)
```

Pass `sheet` to inspect a single worksheet.

```{r sheet-info-one}
excel_sheet_info(file, sheet = "Sheet1")
```

## Columns

`excel_sheet_columns()` returns the names, positions, and inferred dtypes for a
worksheet selection.

```{r sheet-columns}
excel_sheet_columns(file)
```

The same row, column, and dtype options used by `read_excel()` can be used when
inspecting columns.

```{r selected-columns}
excel_sheet_columns(file, columns = c("city", "Year"))
excel_sheet_columns(file, range = "A:C")
```

Set `available = TRUE` to inspect all available columns after applying header
and dtype options, while ignoring `range` and `columns`.

```{r available-columns}
excel_sheet_columns(file, columns = "city", available = TRUE)
```

## Tables and Defined Names

Use `excel_tables()` and `excel_table_info()` to discover named Excel tables.
The bundled example workbook does not contain tables, so these return empty
results for this file.

```{r tables}
excel_tables(file)
excel_table_info(file)
```

When a workbook contains a named Excel table, read it directly with
`read_excel_table()`.

```{r read-table, eval = FALSE}
read_excel_table(file, "Table1", as = "data.frame")
```

Defined names can be inspected with `excel_defined_names()`.

```{r defined-names}
excel_defined_names(file)
```
